A Study on Common Bug Fix Patterns in Rust

### 2022-10-11 sections/discussion.tex
discussion v1 added

@@

In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more value to tree elements closer to the leaves. In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recored their results separate from general results. As this approach is based on an intuition, it does not guarantee achieving optimal clusters. (2) The collected commits are the ones that fix a bug in a previously pushed commit. We do not consider the bugs that occur during a commit and also fixed during the same commit. (3) Our data modelling approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm. Other program data modelling methods and clustering algorithms (e.g. k-means, or SLINK) might output new clusters that our pipeline is unable to find.

The threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality. Similarly, a commit message might contain our target keywords while the commit is not associated with bug fixing changes.

>>>>

In this work, our pipeline successfully mined 19 cross project bug fix patterns. However, our pipeline fails to detect complex pattern due to space limitation and their innate complicated features. A complicated pattern is a pattern that includes many changes in many places and it makes it easier for two datapoints to be dissimilar.

Also, we expected to obtain certain patterns regarding some ertain Rust related features, such as unsafe Rust, but our results did not include those patterns. Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors. That is why, we think there are some patterns that were hidden from our sight.

We believe our results can help researchers to create Rust tools for repair, fault localization, IDE linters, and etc. Also, these results can help program encoding methods to focus on more important aspects of a program and shrink down information more effectively.

@@

In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more values to tree elements closer to the leaves. In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recorded their results separate from general results. As this approach is based on an intuition, it does not guarantee achieving optimal clusters. (2) The collected commits are the ones that fix a bug in a previously pushed commit. We do not consider the bugs that occur during a commit and also fixed during the same commit. (3) Our data modelling approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm. Other program data modelling methods and clustering algorithms (e.g.SLINK) might output new clusters that our pipeline is unable to find.

The threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality. Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.

### 2022-10-11 sections/results.tex
discussion v1 added

As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fix-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST. Here we want to evaluate the effectiveness of our encoding.

The general idea for our evaluation is whether the most important program elements has the high values in our encoding. The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point. We used circle pack figures to visualize each data point. In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Fig ref).

@@

We carried out two manual experiments for evaluating our encoding mechanism. In the first experiment, the analyzer investigates whether the most important element has the largest radius or not. In the second experiment, the analyzer investigates whether the most important element is included in three largest circles or not. The analyzer should give a binary answer to these questions. Both authors carry out these experiments independently. At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness. The final results are TBD\% and TBD\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.

@@

For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option<MediaQueryList>, and modified the other parts of the source code, accordingly.

@@

\begin{table}[]
\begin{tabular}{l|l|r|r}
\textbf{Catgegory} & \textbf{ID} & \textbf{\# Instances} & \textbf{\# Projects} \\
\hline
\multirow{2}{*}{Clone}                                         & BC.clone.ref & 4 & 3 \\
& BC.clone.drop & 83 & 10   
\\\cline{1-4}
\multirow{3}{*}{Ref and Deref} & BC.ref.add & 17 & 7 \\
& BC.ref.drop & 145 & 14 \\
& BC.deref.to\_string  & 7 & 4                                            \\\cline{1-4}
\multirow{2}{*}{Mut} & BC.mut.add  & 11 & 6 \\
& BC.mut.drop  & 42 & 15                               \\\cline{1-4}
Vector & BC.vec.slice  & 8 & 6                                    \\\cline{1-4}
\multirow{1}{*}{Lifetime}      
& BC.lifetime.static  & 8 & 6
\\
\end{tabular}
\caption{\label{table:bc}Bug Fix Pattern Categorization (BC-Related Patterns)}
\end{table}

@@

@@

\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints. Clippy detects whether a variable clone is necessary and could we have used the variable itself or not. This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.

@@

To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git\_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope. Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug. \\

@@

\noindent \textbf{Description:} This is a change that in all the instances has been detected by Clippy. The change applies on an object prior to calling $to\_string()$ on that object. The change might add or remove a dereference, as shown above. In the addition case, the reason behind the change is that the object a reference type of $T$, while $T$ directly implements $to\_string()$. It is faster and more performance efficient to dereference the object and then call $to\_string()$ on it. The deletion case indicates a redundant dereference, thus it is removed.

@@

\noindent \textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword. In the observed instances, this pattern was not seen in refactoring changes, the modification happened alongside a change in program behaviour.

>>>>

As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fix-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements). Here we want to evaluate the effectiveness of our encoding.

The general idea for our evaluation is whether the most important program elements has the high values in our encoding. The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point. We used circle pack figures to visualize each data point \cite{collins2003circle}. For this purpose, we used packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Similar to Figure~\ref{fig:essence}).

@@

We carried out two manual experiments for evaluating our encoding mechanism. In the first experiment, the analyzer investigates whether the most important element has the largest radius or not. In the second experiment, the analyzer investigates whether the most important element is included in three largest circles or not. The analyzer should give a binary answer to these questions. Both authors carry out these experiments independently. At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness. The final results are 55\% and 78\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.

@@

For instance, in the Rust project swc-project/swc (A fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code in a way that it accounts for empty @media queries \footnote{\url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}, so, as shown in the code snippet above, they changed type MediaQueryList to Option\verb+<+MediaQueryList\verb+>+, and modified the other parts of the source code, accordingly.

@@

\begin{table}[]
\begin{tabular}{l|l|r|r}
\textbf{Catgegory} & \textbf{ID} & \textbf{\# Instances} & \textbf{\# Projects} \\
\hline
\multirow{2}{*}{Clone}                                         & BC.clone.drop & 83 & 10 \\
& BC.clone.ref & 4 & 3   
\\\cline{1-4}
\multirow{3}{*}{Ref and Deref} & BC.ref.add & 17 & 7 \\
& BC.ref.drop & 145 & 14 \\
& BC.deref.to\_string  & 7 & 4                                            \\\cline{1-4}
\multirow{2}{*}{Mut} & BC.mut.add  & 11 & 6 \\
& BC.mut.drop  & 42 & 15                               \\\cline{1-4}
Vector & BC.vec.slice  & 8 & 6                                    \\\cline{1-4}
\multirow{1}{*}{Lifetime}      
& BC.lifetime.static  & 8 & 6
\\
\end{tabular}
\caption{\label{table:bc}Bug Fix Pattern Categorization (BC-Related Patterns)}
\end{table}

@@

\noindent \textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints. Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes. There are more than 500 lints included in Clippy. With regards to this pattern, Clippy detects whether a variable clone is necessary and could we have used the variable itself or not. This change does not modify the program behaviour, however it can boost the performance significantly if we repeatedly cloning large variables.

@@

@@

To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git\_branches.rs}}, the developer borrows the value of \verb+repo_dir+ to keep the ownership of the object in the current scope. Now this object can then be used in another function invocation (\verb+remove_dir_all(repo_dir)+) which fixes a bug. \\

@@

\noindent \textbf{Description:} This is a change that in all the instances has been detected by Clippy. The change applies on an object prior to calling \verb+to_string()+ on that object. The change might add or remove a dereference, as shown above. In the addition case, the reason behind the change is that the object a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+. It is faster and more performance efficient to dereference the object and then call \verb+to_string()+ on it. The deletion case indicates a redundant dereference, thus it is removed.

@@

\noindent \textbf{Description:} This change happens when the programmer needs to mutate a variable which is immutable, hence they need to add the mut keyword. In the observed instances, this pattern was not seen in refactoring changes, the modification happened alongside a change in program behaviour. \\
